Stream graphes multicouches

### 2019-07-25 tex/article/main.tex
update article

\documentclass{svproc}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

@@

This formalism can be extended to capture diversity in the data: links can be oriented, weighted, labelled, \textit{etc}. This different notions has been gathered in a new formalism, the multilayer graphs~\cite{mlkiv}. Furthermore,

\section{The multilayer stream graph}
%
\subsection{Preliminaries: multilayer graphs and stream graphs}
%

\subsubsection{Multilayer graphs}

\subsubsection{Stream graphs}
\subsection{Definition of the multilayer stream graph}
%
We define a multilayer stream graph (MSG) as a tuple $S_M = (T,T_M,V,W_M,E_M,{\cal L})$, such that $T$ and $V$ are, respectively, a time interval and a set of nodes, just like for stream graphs \cite{stream}.

@@

   $E_M \subseteq T\times V_M \otimes V_M$ are the links appearing with respect to time,  $V_M = \{ (u,\alpha), u\in V, \alpha \in L\}$. The elements of $V_M$ are named {\em node-layer}. A link cannot appear outside the time of existence of the two node-layers.

\subsection{Notions associated}
%
\subsubsection{Basic notions}
%
\subsubsection{Centralities}
%
\subsection{Implementation and complexity}
%
\subsection{Autonomous Systems}
%

\section{Datasets and results}
%
\subsection{Two datasets}
%
\subsubsection{Classes in an highschool: a social multilayer stream graph}
%
\subsubsection{Civil airlines in US: an application to transportation}
%
\subsection{Analysis}
%

@@

>>>>

\documentclass{svproc}
\usepackage{amsmath}
\usepackage{cleveref}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

%tikz 
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}
\usetikzlibrary{decorations.pathmorphing, positioning}
\definecolor{echoreg}{HTML}{2cb1e1}
\definecolor{echodrk}{HTML}{0099cc}
\tikzstyle{mybox} = [text=black, very thick,
    rectangle, rounded corners, inner sep=10pt, inner ysep=20pt]
\tikzstyle{fancytitle} =[text=black]
\newcommand{\yslant}{0.5}
\newcommand{\xslant}{-0.6}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TIKZ STYLES
\tikzstyle{every node}=[font=\scriptsize]

\tikzset{RootStyle/.style = {
						shape          = circle,
			      draw           = red!50!black!50,
			      thick,
			      top color      = white,
			      bottom color   = red!50!black!20,
			      text           = black,
			      inner sep      = .2pt,
			      outer sep      = 0pt,
			      minimum size   = 2.5 mm}
						}

\tikzset{VertexStyle/.style = {
						shape          = circle,
			      draw           = black!50,
			      top color      = white,
			      bottom color   = black!20,
			      text           = black,
			      inner sep      = .2pt,
			      outer sep      = 0pt,
			      minimum size   = 8 mm}
						}

\tikzset{EdgeStyle/.style   = {thick,-}}

\tikzset{LabelStyle/.style =   {
				  text           = black,
				  inner sep      = .2pt,
				  outer sep      = 1pt,
				  font           =\scriptsize,
				  minimum size   = 2.15 mm}
					}

\pgfplotsset{compat=1.15}

\newcommand\overmat[3]{%
  \makebox[0pt][l]{$\smash{\color{#3}\overbrace{\phantom{%
    \begin{matrix}#2\end{matrix}}}^{\text{#1}}}$}#2}
\newcommand\undermat[3]{%
  \makebox[0pt][l]{$\smash{\color{#3}\underbrace{\phantom{%
    \begin{matrix}#2\end{matrix}}}_{\text{#1}}}$}#2}
\newcommand\partialphantom{\vphantom{\frac{\partial e_{P,M}}{\partial w_{1,1}}}}

@@

This formalism has been extended to capture diversity in the data: links can be oriented, weighted, labelled, \textit{etc}. The nodes can have different labels too. The nodes and the links can change with the time. Several new formalisms have been created to deal with this new aspects. We focus on two of them: the multilayer graphs~\cite{mlkiv} to deal with non-trivial structures and the stream graphs~\cite{stream} to deal with time-dependant interactions. We use this two formalisms and the theories associated to describe time-dependant set of interactions with complex structures.

In the first part, the multilayer stream graph is constructed with its theoretical definitions. The we give example of use with to real datasets.

\section{The multilayer stream graph}
%

\subsection{Preliminaries: multilayer graphs and stream graphs}
%
An graph is a tuple $G=(V,E)$ composed by a set of nodes $V$ and a set of edges $E\subseteq V \otimes V$, where each edge is an unordered pair of two different nodes. A directed graph has a set of directed edges $E\subseteq V\times V$, where $(u,v) \neq (v,u)$.

To illustrate the definitions of all the graphs, we will use the same example: a population of monkeys $F_1, F_2, M_1, M_2$ (two females and two males). In the \cref{fig:graphsimple}, we put a link between two animals if they have been in contact.

\begin{figure}[h]
			\begin{center}
					\begin{tikzpicture}[xscale=2,yscale=2,auto]
							\draw (1,0) node[VertexStyle]   (A) { $F_1$ };
							\draw (0,0) node[VertexStyle] (B) { $F_2$ };
							\draw (0,1) node[VertexStyle]   (C) { $M_1$ };
							\draw (1,1) node[VertexStyle]   (D) { $M_2$ };

							\draw[EdgeStyle]				(B) to node[LabelStyle]{}	(A);
							\draw[EdgeStyle,bend left=40,looseness=1.1]    	(C) to node[LabelStyle]{}   	(A);
							\draw[EdgeStyle] 	(C) to node[LabelStyle]{}     	(B);
							\draw[EdgeStyle]  	(C) to node[LabelStyle]{} 		(D);

					\end{tikzpicture}
			\end{center}
	\caption{An example of undirected graph. The nodes $V=(F_1, F_2, M_1, M_2)$ are monkeys, and the undirected edges $E=(M_1,M_2),(M_1,F_1),(M_1,F_2),(F_1,F_2)$ represent the fact that they have been in contact.}
	\label{fig:graphsimple}
\end{figure}

The {\bf degree} of a node $v\in V$ $d(v)$ is the number of edges in which $v$ is appears: $d(v)=|\{(u,w) \in E \subseteq V \otimes V| u=v \cup w=v\}|$.

\subsubsection{Multilayer graphs}
%
Imagine now that we want to describe in a more precise way the type of interaction between the monkeys. For instance, the monkey can meet in different places, like the mountain or the plain. The can have face to face relations to collaborate or to fight.

A {\em multilayer graph}~\cite{mlkiv,mldd} is a set $M=(V_M,E_M,V,{\cal L})$.\\
${\cal L}$ is the {\em structure}: it is a finite set of $d$ different sets, named {\em aspects}: ${\cal L} = L_1, \dots, L_d$. Each aspect $L_i$ contains elements $l_i^1,\dots l_i^{n_i}$ which are named {\em elementary layers}. A {\em layer} $\alpha$ is then a combination of elementary layers from each aspect: $\alpha \in L=L_1\times \dots \times L_k$. \\
$V$ is the set of nodes and each of them can be present or not on the different layers. A node on a layer will be called a {\em node-layer}. The set of nodes-layers is $V_M \subseteq V \times L$. Each node-layer can be linked to another with undirected ($E_M\subseteq V_M \otimes V_M$)or directed ($E_M \subseteq V_M \times V_M$) edges.

In the monkeys example \cref{multimonkeys}, the nodes are still the monkeys. The structure is made of two aspects: the place and the type of interaction. A layer can be for instance (mountain,collaboration). They are different types of links, depending on which layer are the nodes.

\begin{equation}
\begin{array}{rcl}
  {\cal L}&=&\{\text{place},\text{type of relationship}\} \\
  \text{place} &=& \{\text{mountain, plane}\}\\
 \text{type of relationship} &=& \{ \text{fight,collaboration}\}
\end{array}
\end{equation}
%0.58

\begin{figure}[h]
\begin{center}
\begin{tikzpicture}[scale=0.38,every node/.style={minimum size=1cm},on grid]

	\node [mybox, scale=1.0] at (10.5, 2) (box){%
		\begin{minipage}{0.6\textwidth}

    	\end{minipage}
	};

	% etage 2
	\begin{scope}[
		yshift=-210,
		every node/.append style={yslant=\yslant,xslant=\xslant},
		yslant=\yslant,xslant=\xslant
	] 
		%\draw[black, dashed, thin] (0,0) rectangle (7,7); 
		\fill[olive,fill opacity=.75] (0,0) rectangle (7,7);

		\fill[orange,fill opacity=.75] (10,0) rectangle (17,7);
		%\draw[black, dashed, thin] (10,0) rectangle (17,7);

		\draw[fill=echoreg]  %foret, collabo
			(5,3) node(111){} circle (.1) %A
			(2,3) circle (.1) %B
			(2,6) circle (.1); %C
			%(5,5) circle (.1); %D

		\draw[fill=echoreg]  %plaine, collabo
			(15,3) node(111){} circle (.1) %A
			%(12,2) circle (.1) %B
			(12,6) circle (.1) %C
			(15,6) circle (.1); %D

		\draw[ thin, color=echodrk]%foret collab
			(2,5.9) to (2,3.1); %C->B
		\draw[ thin, color=echodrk]
			(2.1,3) to (4.9,3);%B->A

		\draw[thin, color=echodrk]%plaine collab
			(12,6) to (15,3);%C->A
		\draw[thin, color=echodrk]
			(12,6) to (15,6);

		\fill[black]
			(0.5,1.5) node[right, scale=.7] {\bf \large Mountain,} 	
			(0.5,0.5) node[right, scale=.7] {\bf \large Collaboration}
			(5.1,2.9) node[right,scale=.9]{$\mathbf{F_1}$}
			(1.9,2.9) node[left,scale=.9]{$\mathbf{F_2}$}
			(2,6.2) node[left,scale=.9]{$\mathbf{M_1}$};
			%(5.2,5.1) node[right,scale=.7]{\bf D};

		\fill[black]
			(10.5,1.5) node[right, scale=.7] {\bf \large Plain,}
			(10.5,0.5) node[right, scale=.7] {\bf \large Collaboration}
			(15.1,2.9) node[right,scale=.9]{$\mathbf{F_1}$}
			%(11.9,1.9) node[left,scale=.7]{\bf B}
			(12,6.2) node[left,scale=.9]{$\mathbf{M_1}$}
			(15.2,6.1) node[right,scale=.9]{$\mathbf{M_2}$};

	\end{scope}

	% Interlayer crossconnections
	% vertical
	\draw[thick,  dashed, decorate] (3.2, 4.7) to (3.2, -2.8);%A
	\draw[thick, dashed, decorate] (.2,3.1) to (.2,-4.3);%B
	\draw[thick,  dashed, decorate] (-1.6, 5.2) to (-1.6, -2.1);%C

1	\draw[thick,  dashed, decorate] (13.2,9.7) to (13.2, 2.2);%A
	\draw[thick, dashed, decorate] (11.4,11.7) to (11.4,4.3);%D
	\draw[thick,  dashed, decorate] (8.4, 10.2) to (8.4, 2.8);%c

	%horizontal

	\draw[thick, dashed, decorate] (-1.6,-2.2) to[bend left] (8.4,2.8);
	\draw[thick, dashed, decorate] (3.2,-2.8) to[bend left] (13.2,2.2);

	% etage 1
	\begin{scope}[
		yshift=0,
		every node/.append style={yslant=\yslant,xslant=\xslant},
		yslant=\yslant,xslant=\xslant
	]
		\fill[olive,fill opacity=.85] (0,0) rectangle (7,7); 
		%\draw[black, dashed, thin] (0,0) rectangle (7,7);

		\fill[orange,fill opacity=.85] (10,0) rectangle (17,7);
		%\draw[black, dashed, thin] (10,0) rectangle (17,7);

		\draw [fill=red]
			(5,3) node(111){} circle (.1)%A %foret, combat
			(2,3) circle (.1)%B
			(2,6) circle (.1)%C
			(5,6) circle (.1);%D

		\draw[fill=red]  
			(15,3) node(111){} circle (.1) %A % plaine combat
			%(12,2) circle (.1) %B
			(12,6) circle (.1) %C
			(15,6) circle (.1); %D

		\draw[thin, color=red]%foret combat
			(2,3.1) to (2,5.9);%B->C

		\draw[thin, color=red]%plaine combat
			(12,6) to (15,6);%C->D

		\fill[black]
			(0,0.5) node[right, scale=0.7] {\bf \large Mountain, Fight}
			(5.1,2.9) node[right,scale=0.9]{$\mathbf{F_1}$}
			(1.9,2.9) node[left,scale=0.9]{$\mathbf{F_2}$}
			(1.9,6) node[left,scale=0.9]{$\mathbf{M_1}$}
			(5.2,6.1) node[right,scale=0.9]{$\mathbf{M_2}$};

		\fill[black]
			(10.5,0.5) node[right, scale=.7] {\bf \large Plain, Fight}
			(15.1,2.9) node[right,scale=.9]{$\mathbf{F_1}$}
			%(11.9,1.9) node[left,scale=.7]{\bf B}
			(12,6.2) node[left,scale=.9]{$\mathbf{M_1}$}
			(15.2,6.1) node[right,scale=.9]{$\mathbf{M_2}$};

	\end{scope}

	%interlayer
	\draw[thick, dashed, decorate] (-1.6,5.2) to[bend left] (8.4,10.2);%C
	\draw[thick, dashed, decorate] (3.2,4.7) to[bend left] (13.2,9.7);%A
	\draw[thick, dashed, decorate] (1.5,6.8) to[bend left] (11.4,11.7);%D
\end{tikzpicture}
\caption{The multilayer graph of relationships between monkeys. Red link between $M_1$ and $F_2$ means that the two monkeys have been fighting in the forest. The dotted links means that the nodes-layers are from the same nodes.}
\label{multimonkeys}
\end{center}
\end{figure}

\subsubsection{Stream graphs}
From the beginning, we added a link between two monkeys if they have been in contact, but we do not know anything about the times of existence of these contacts. We can use the stream graph to model this.

A {\em stream graph}~\cite{stream} is a set of four sets $S=(T,W,V,E)$.\\
$T$ is the time interval of study, $V$ is the set of nodes. The {\em time-nodes} set $W \subseteq T \times V$ describes the existence of nodes depending on the time: $(t,v) \in W$ means that the node $v$ appears at $t$. $E_M \subseteq T \times V \times V$ contain all the links and their time of existence.

Given nodes $u$ and $v$, we call $T_u$ the set of times at which $u$ appears, and $T_uv$ the set of time at which the link $(u,v)$ appears.

\begin{equation}
\begin{array}{rcl}
T_u&=&\{t, (t,u) \in W \}\\
T_{uv}&=&\{t, (t,uv)\in E \} 
\end{array}
\end{equation}

In the \cref{exmonkeystream}, we give an example of stream graph for the monkey population. We look at a place (let's say the forest) and we put a link between two monkeys when they interact.

\begin{figure}
\centering
	\includegraphics[width=0.5\textwidth]{img/ext.pdf}
	\caption{Example of stream graph describing interactions between monkeys in the forest. $F_1$ arrives at $t=3$, and interacts with $F_2$ at $t \in[4,5]\cup [7,8]$. $F_2$ and $M_1$ interact between $t=2$ and $4$.}
	\label{exmonkeystream}
\end{figure}

The stream graph model does not require, as for other temporal graph models, to discretize the time. Several notions have been designed on \cite{stream} from the model of the graph notions. Among them, we can find the number of nodes, the density, the uniformity, the compacity , the degree, the betweenness centrality, path and distances.

\subsection{Definition of the multilayer stream graph}
%
We now gather the notion of multilayer graphs and stream graphs to capture both structure and temporal subtleties.

We define a multilayer stream graph (MSG) as a tuple $S_M = (T,T_M,V,W_M,E_M,{\cal L})$, such that $T$ and $V$ are, respectively, a time interval and a set of nodes, just like for stream graphs \cite{stream}.

@@

   $E_M \subseteq T\times V_M \otimes V_M$ are the links appearing with respect to time,  $V_M = \{ (u,\alpha), u\in V, \alpha \in L\}$. The elements of $V_M$ are named {\em node-layer}. A link can not appear outside the time of existence of the two node-layers.

   \begin{figure}[h]
   	\includegraphics[width=0.8\textwidth]{img/monkeymls.pdf}
   	\caption{An example of the 2-layers stream graph of monkeys, collaborating in mountain or in plain. $F_1$ is in [mountain, collaboration] between $t=0$ and $t=4$ and then goes to [plain,collaboration]. It collaborates with $F_2$ at $t\in [1,2]\cup[3,4]$: $\{(t,F_1,[\text{mountain,collaboration}],F_2,[\text{mountain,collaboration}]),t\in [1,2]\cup[3,4]\} \subset E_M$.}
   \end{figure}

\subsection{Notions associated}
%
\subsubsection{Basic notions}
%

\subsubsection{Centralities}
%
\subsection{Implementation and complexity}
%
\subsection{Autonomous Systems}
%

\section{Datasets and results}
%
\subsection{Two datasets}
%
\subsubsection{Classes in an high school: a social multilayer stream graph}
%
\subsubsection{Civil airlines in US: an application to transportation}
%
\subsection{Analysis}
%

@@

\bibitem{mldd}
De Domenico M., Sol\'e-Ribalta A., Cozzo E., Kivel\"a M., Moreno Y., Porter M.A., G\'omez S., Arenas A.: Mathematical Formulation of Multilayer Networks.(2013). In: Phys. Rev. X. doi:10.1103/PhysRevX.3.041022

### 2019-07-25 tex/rapport de stage/rapport.tex
update article

 $V$ est l'ensemble des {\em n\oe{}uds}, l'ensemble $V_M \subseteq V\otimes L$ les {\em n\oe{}uds-couches}. Enfin $E_M \subseteq V_M \times V_M $ est l'ensemble des {\em arêtes}, une arête pouvant relier deux n\oe{}uds-couches entre eux.

>>>>

 $V$ est l'ensemble des {\em n\oe{}uds}, l'ensemble $V_M \subseteq V\times L$ les {\em n\oe{}uds-couches}. Enfin $E_M \subseteq V_M \otimes V_M $ est l'ensemble des {\em arêtes}, une arête pouvant relier deux n\oe{}uds-couches entre eux.
